---
title: "Bootstrapping SVM global genes ML"
output: html_notebook
---

## 1.

```{r}
set.seed(30921)  # sets anchor for random number generation -> reproducible code
res.sampling.names <- sample(x = res.cellNames, size = 426, replace = FALSE) # random sample of 426 cells
bias.corrected.res <- subset(ag.training, cells = res.sampling.names, invert = TRUE)
res.testing <- subset(ag.training, cells = res.sampling.names, invert = FALSE)

set.seed(31021)
testing.names <- sample(x = colnames(bias.corrected.res), size = 104, replace = FALSE) # random sample of 104 cells
testing.cells <- subset(bias.corrected.res, cells = testing.names)
training.cells <- subset(bias.corrected.res, cells = testing.names, invert = TRUE)

training.genes <- rownames(training.cells)
training.cells <- training.cells %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = training.genes) %>%
  RunPCA(features = VariableFeatures(object = training.cells)) %>%
  RunUMAP(dims = 1:30)

all <- getFeatureSpace(object = training.cells, pvar = "phenotype", correction = "fdr", sig = 1, reduction = 'pca')

all <- trainModel(all)
all.probabilities <- get_probabilities(all) # probabilities of each classification
all.summary <- get_scpred(all) # summary table

testing <- merge(testing.cells, res.testing)
testing <- NormalizeData(testing)
all.test <- scPredict(testing, all)

all.verify <- crossTab(all.test, "phenotype", "scpred_prediction")
all.summary
all.verify
```
ML model with global genes summary (boostrap 2)
· 89.43% success based on 474/530 testing data points
· res: ROC 0.992, sensitivity 0.959, specificity 0.963
· vuln: ROC 0.992, sensitivity 0.963, specificity 0.959


## 2.

```{r}
set.seed(31121)  # sets anchor for random number generation -> reproducible code
res.sampling.names <- sample(x = res.cellNames, size = 426, replace = FALSE) # random sample of 426 cells
bias.corrected.res <- subset(ag.training, cells = res.sampling.names, invert = TRUE)
res.testing <- subset(ag.training, cells = res.sampling.names, invert = FALSE)

set.seed(31221)
testing.names <- sample(x = colnames(bias.corrected.res), size = 104, replace = FALSE) # random sample of 104 cells
testing.cells <- subset(bias.corrected.res, cells = testing.names)
training.cells <- subset(bias.corrected.res, cells = testing.names, invert = TRUE)

training.genes <- rownames(training.cells)
training.cells <- training.cells %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = training.genes) %>%
  RunPCA(features = VariableFeatures(object = training.cells)) %>%
  RunUMAP(dims = 1:30)

all <- getFeatureSpace(object = training.cells, pvar = "phenotype", correction = "fdr", sig = 1, reduction = 'pca')

all <- trainModel(all)
all.probabilities <- get_probabilities(all) # probabilities of each classification
all.summary <- get_scpred(all) # summary table

testing <- merge(testing.cells, res.testing)
testing <- NormalizeData(testing)
all.test <- scPredict(testing, all)

all.verify <- crossTab(all.test, "phenotype", "scpred_prediction")
all.summary
all.verify
```
ML model with global genes summary (boostrap 3)
· 89.06% success based on 472/530 testing data points
· res: ROC 0.992, sensitivity 0.963, specificity 0.957
· vuln: ROC 0.992, sensitivity 0.957, specificity 0.963


## 4.

```{r}
set.seed(31321)  # sets anchor for random number generation -> reproducible code
res.sampling.names <- sample(x = res.cellNames, size = 426, replace = FALSE) # random sample of 426 cells
bias.corrected.res <- subset(ag.training, cells = res.sampling.names, invert = TRUE)
res.testing <- subset(ag.training, cells = res.sampling.names, invert = FALSE)

set.seed(31421)
testing.names <- sample(x = colnames(bias.corrected.res), size = 104, replace = FALSE) # random sample of 104 cells
testing.cells <- subset(bias.corrected.res, cells = testing.names)
training.cells <- subset(bias.corrected.res, cells = testing.names, invert = TRUE)

training.genes <- rownames(training.cells)
training.cells <- training.cells %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = training.genes) %>%
  RunPCA(features = VariableFeatures(object = training.cells)) %>%
  RunUMAP(dims = 1:30)

all <- getFeatureSpace(object = training.cells, pvar = "phenotype", correction = "fdr", sig = 1, reduction = 'pca')

all <- trainModel(all)
all.probabilities <- get_probabilities(all) # probabilities of each classification
all.summary <- get_scpred(all) # summary table

testing <- merge(testing.cells, res.testing)
testing <- NormalizeData(testing)
all.test <- scPredict(testing, all)

all.verify <- crossTab(all.test, "phenotype", "scpred_prediction")
all.summary
all.verify
```
ML model with global genes summary (boostrap 4)
· 86.23% success based on 457/530 testing data points
· res: ROC 0.99, sensitivity 0.946, specificity 0.955
· vuln: ROC 0.99, sensitivity 0.955, specificity 0.946


## 5.

```{r}
set.seed(31521)  # sets anchor for random number generation -> reproducible code
res.sampling.names <- sample(x = res.cellNames, size = 426, replace = FALSE) # random sample of 426 cells
bias.corrected.res <- subset(ag.training, cells = res.sampling.names, invert = TRUE)
res.testing <- subset(ag.training, cells = res.sampling.names, invert = FALSE)

set.seed(31621)
testing.names <- sample(x = colnames(bias.corrected.res), size = 104, replace = FALSE) # random sample of 104 cells
testing.cells <- subset(bias.corrected.res, cells = testing.names)
training.cells <- subset(bias.corrected.res, cells = testing.names, invert = TRUE)

training.genes <- rownames(training.cells)
training.cells <- training.cells %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = training.genes) %>%
  RunPCA(features = VariableFeatures(object = training.cells)) %>%
  RunUMAP(dims = 1:30)

all <- getFeatureSpace(object = training.cells, pvar = "phenotype", correction = "fdr", sig = 1, reduction = 'pca')

all <- trainModel(all)
all.probabilities <- get_probabilities(all) # probabilities of each classification
all.summary <- get_scpred(all) # summary table

testing <- merge(testing.cells, res.testing)
testing <- NormalizeData(testing)
all.test <- scPredict(testing, all)

all.verify <- crossTab(all.test, "phenotype", "scpred_prediction")
all.summary
all.verify
```

ML model with global genes summary (boostrap 5)
· 87.74% success based on 465/530 testing data points
· res: ROC 0.992, sensitivity 0.958, specificity 0.958
· vuln: ROC 0.992, sensitivity 0.958, specificity 0.958

## Final>

Across 5 bootstrap replicates (resampling training and testing data):
· Average success rate:     88.303%
· Average ROC:              0.9916
· Average res sensitivity:  0.9566
· Average res selectivity:  0.9586
· (vuln sens and selectivity are the same as res but reversed)

