---
title: Training data prep
output: html_notebook
---

## 0.
Data is from the transcriptomes of 6 mouse spinal cords, sequenced in Blum 21. Download data from:
路 Alpha:    http://spinalcordatlas.org/assets/downloads/alphamns.h5ad
路 Gamma:    http://spinalcordatlas.org/assets/downloads/gammamns.h5ad
路 Visceral: http://spinalcordatlas.org/assets/downloads/visceralmns.h5ad
路 Whole dataset: http://spinalcordatlas.org/assets/downloads/allexpsctl.h5ad


## 1. 
Import libraries, create directories for saving images

```{r}
library(Seurat)
library(zellkonverter)
library(ggplot2)
library(SummarizedExperiment)
library(SingleCellExperiment)
library(magrittr)

rds.save.dir <- "/Users/xuanzhao/Documents/R/PredictResistance/"
graph.save.dir <- "/Users/xuanzhao/Documents/R/PredictResistance/graphs/"
```

## 2.
Load data (h5ad files) as objects of class SingleCellExperiment

```{r}
alpha.data <- readH5AD(file = "~/Documents/R/GSE161621_filtered_data/alphamns.h5ad")
gamma.data <- readH5AD(file = "~/Documents/R/GSE161621_filtered_data/gammamns.h5ad")

```

## 3.
Convert objects to class Seurat

```{r}
a.counts <- assay(alpha.data, "X")
a.libsizes <- colSums(a.counts)
a.size.factors <- a.libsizes/mean(a.libsizes)
logcounts(alpha.data) <- log2(t(t(a.counts)/a.size.factors) + 1)            # calculate + add log-transformed normalized expression values from the count matrix 
logcounts(alpha.data) = as.matrix(logcounts(alpha.data))                      # convert logcounts sparse matrix to matrix
alpha <- as.Seurat(alpha.data, counts = "X", data = "logcounts")  # convert to seurat object

g.counts <- assay(gamma.data, "X")
g.libsizes <- colSums(g.counts)
g.size.factors <- g.libsizes/mean(g.libsizes)
logcounts(gamma.data) <- log2(t(t(g.counts)/g.size.factors) + 1)
logcounts(gamma.data) = as.matrix(logcounts(gamma.data))
gamma <- as.Seurat(gamma.data, counts = "X", data = "logcounts")
```

## 4.
Split alpha MNs into fast-firing, fast-fatigue-resistant, and slow-firing (ff, fr, sf)

```{r}
DefaultAssay(alpha) <- 'RNA' # default assay set to RNA

Idents(object = alpha, WhichCells(object = alpha, expression = Prkcd>0 & Sv2a>0 & Chodl==0, slot = 'data')) <- "sf"

Idents(object = alpha, WhichCells(object = alpha, expression = Chodl>0 & Kcnq5>0 & Sv2a==0, slot = 'data')) <- "ff"

Idents(object = alpha, WhichCells(object = alpha, expression = Chodl>0 & Kcnq5==0 & Sv2a==0, slot = 'data')) <- "fr"

alpha <- subset(x = alpha, idents = "SingleCellExperiment", invert = TRUE) # removes ident SingleCellExperiment

```

## 5.
Split gamma MNs into gamma and beta by ident

```{r}

beta.subset <- subset(x = gamma, subset = skeletal_subtype == 'Beta(?) Motor Neurons')
Idents(object = gamma, WhichCells(object = gamma, cells = colnames(beta.subset))) <- 'beta'

gamma.subset <- subset(x = gamma, subset = skeletal_subtype == 'Gamma Motor Neurons')
Idents(object = gamma, WhichCells(object = gamma, cells = colnames(gamma.subset))) <- 'gamma'

```

## 6.
Combine datasets

```{r}
ag.combined <- merge(x = alpha, y = gamma)
all.genes <- rownames(ag.combined)

ag.combined <- ag.combined %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = all.genes) %>%
  RunPCA(features = VariableFeatures(object = ag.combined)) %>%
  RunUMAP(dims = 1:10) # dims = 1:30 instead?

# If RunPCA doesn't work, try entering RunPCA and RunUMAP into the console

```

## 7.
General housekeeping

```{r}

# ----------
# Just to know the amount of data per identity

ag.combined.table <- table(Idents(ag.combined))  # table with number of cells in each identity

# ----------
# Visualizing Kaplan 14 genes and their expression levels in this dataset
# Helps determine whether they would be a good fit for ML feature space confinement

AverageExpression(ag.combined, features = c("Sema3e","Gad2","Kcnj5","Rsph9","Npr3","Vipr2","Ecrg4","Egln3","Mgam","Arhgap9","Dlk1","Mmp9","Il1f9","Trhr","Aldh1a2","Cited1","Hsd17b2","Nts"), slot = 'data')

# Likely gene candidates:
# Sema3e (opposite), Gad2, Kcnj5, Arghap9, Dlk1 (maybe opposite), 

```

## 8.
Pairwise DE on g/s v. ff/fr

```{r}

# g.s.markers <- FindMarkers(ag.combined, ident.1 = "gamma", ident.2 = "sf")
# g.ff.markers <- FindMarkers(ag.combined, ident.1 = "gamma", ident.2 = "ff")
# g.fr.markers <- FindMarkers(ag.combined, ident.1 = "gamma", ident.2 = "fr")
# ff.fr.markers <- FindMarkers(ag.combined, ident.1 = "ff", ident.2 = "fr")
# ff.s.markers <- FindMarkers(ag.combined, ident.1 = "ff", ident.2 = "sf")
# fr.s.markers <- FindMarkers(ag.combined, ident.1 = "fr", ident.2 = "sf")

```



