---
title: Iteration 1
output: html_notebook
---

· PCA of global genes used in feature space: no DE to find discriminant genes between res/vuln

## 1.
Random subset of 652 resistant neurons to balance training data (reduce bias). 

Vuln:    652    
Res:     652    + 426 'res' data points left over
Total:   1304

```{r}
set.seed(30821)  # sets anchor for random number generation -> reproducible code
res.sampling.names <- sample(x = res.cellNames, size = 426, replace = FALSE) # random sample of 426 cells
bias.corrected.res <- subset(ag.training, cells = res.sampling.names, invert = TRUE)
res.testing <- subset(ag.training, cells = res.sampling.names, invert = FALSE)
```

## 2.
Randomly subset off 104 cells for testing ML accuracy.

Training: 1200
Testing: 530

```{r}
set.seed(30721)
testing.names <- sample(x = colnames(bias.corrected.res), size = 104, replace = FALSE) # random sample of 104 cells
testing.cells <- subset(bias.corrected.res, cells = testing.names)
training.cells <- subset(bias.corrected.res, cells = testing.names, invert = TRUE)

training.genes <- rownames(training.cells)
training.cells <- training.cells %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = training.genes) %>%
  RunPCA(features = VariableFeatures(object = training.cells)) %>%
  RunUMAP(dims = 1:30)

```

## 3.
Create feature space + train ML

```{r}
all <- getFeatureSpace(object = training.cells, pvar = "phenotype", correction = "fdr", sig = 1, reduction = 'pca')

all <- trainModel(all)
all.probabilities <- get_probabilities(all) # probabilities of each classification
all.summary <- get_scpred(all) # summary table
```

## 4.
Testing model: % accuracy with a known-value test set

```{r}
testing <- merge(testing.cells, res.testing)
testing <- NormalizeData(testing)
all.test <- scPredict(testing, all)

all.verify <- crossTab(all.test, "phenotype", "scpred_prediction")
# 472/530 = 89.06% success
# columns = each type of classified neuron
# rows = where each neuron was classified
```

## 5.
ML model with global genes summary (boostrap 1)
· 89.06% success based on 472/530 testing data points
· res: ROC 0.992, sensitivity 0.957, specificity 0.96
· vuln: ROC 0.992, sensitivity 0.96, specificity 0.957

Iterating across 5 bootstrap replicates (resampling training and testing data):
· Average success rate:     88.303%
· Average ROC:              0.9916
· Average res sensitivity:  0.9566
· Average res selectivity:  0.9586
· Average vuln sensitivity: 0.9586
· Average vuln sensitivity: 0.9566


