---
title: "ML Prediction (SVM)"
output: html_notebook
---

Initial input: a preprocessed Seurat object ag.combined with all training data

## 1.
Import packages

```{r}
library(scPred)
```

## 2.
Load predicting data (query) and convert to a Seurat object

```{r}
visceral.data <- readH5AD(file = "~/Documents/R/GSE161621_filtered_data/visceralmns.h5ad")
v.counts <- assay(visceral.data, "X")
v.libsizes <- colSums(v.counts)
v.size.factors <- v.libsizes/mean(v.libsizes)
logcounts(visceral.data) <- log2(t(t(v.counts)/v.size.factors) + 1)
logcounts(visceral.data) = as.matrix(logcounts(visceral.data))
visceral <- as.Seurat(visceral.data, counts = "X", data = "logcounts")
```

## 3.
Add resistant/vulnerable metadata to training data

```{r}
# pbmc$CellType <- Idents(pbmc)

# Create 'phenotype' dataframe (# of rows = # of rows in ag.combined)
phenotype <- data.frame(rownames(ag.combined@meta.data), row.names = rownames(ag.combined@meta.data))
colnames(phenotype) <- "p"

# Subsets of ff, fr, sf, b, and g data
ff.subset <- subset(ag.combined, idents = "ff")
fr.subset <- subset(ag.combined, idents = "fr")
sf.subset <- subset(ag.combined, idents = "sf")
# beta.subset
# gamma.subset

ff.cellNames <- rownames(ff.subset@meta.data)
fr.cellNames <- rownames(fr.subset@meta.data)
vuln.cellNames <- c(ff.cellNames, fr.cellNames)

sf.cellNames <- rownames(sf.subset@meta.data)
g.cellNames <- rownames(gamma.subset@meta.data)
res.cellNames <- c(sf.cellNames, g.cellNames)

b.cellNames <- rownames(beta.subset@meta.data)

# Iteratively changes dataframe
phenotype$p[phenotype$p %in% vuln.cellNames] <- "vuln"
phenotype$p[phenotype$p %in% res.cellNames] <-  "res"
phenotype$p[phenotype$p %in% b.cellNames] <-  "beta_predict"

# Adds dataframe into ag.combined@meta.data - Can now be accessed by ag.combined$phenotype
ag.combined <- AddMetaData(object = ag.combined, metadata = phenotype, col.name = "phenotype")
```

## 4.
Take out beta cells from training data

```{r}
ag.training <- subset(ag.combined, idents = 'beta', invert = TRUE)
```

## 5.
Random subset of 652 resistant neurons to balance training data (reduce bias). 

Vuln:    652    
Res:     652    + 426 'res' data points left over
Total:   1304

```{r}
set.seed(30821)  # sets anchor for random number generation -> reproducible code
res.sampling.names <- sample(x = res.cellNames, size = 426, replace = FALSE) # random sample of 426 cells
bias.corrected.res <- subset(ag.training, cells = res.sampling.names, invert = TRUE)
res.testing <- subset(ag.training, cells = res.sampling.names, invert = FALSE)
```

## 6.
Randomly subset off 104 cells for testing ML accuracy.

Training: 1200
Testing: 530

```{r}
set.seed(30721)
testing.names <- sample(x = colnames(bias.corrected.res), size = 104, replace = FALSE) # random sample of 104 cells
testing.cells <- subset(bias.corrected.res, cells = testing.names)
training.cells <- subset(bias.corrected.res, cells = testing.names, invert = TRUE)

training.genes <- rownames(training.cells)
training.cells <- training.cells %>%
  NormalizeData() %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>%
  ScaleData(features = training.genes) %>%
  RunPCA(features = VariableFeatures(object = training.cells)) %>%
  RunUMAP(dims = 1:30)

```

## 7.
Create feature space + train ML

```{r}
all <- getFeatureSpace(object = training.cells, pvar = "phenotype", correction = "fdr", sig = 1, reduction = 'pca')

all <- trainModel(all)
all.probabilities <- get_probabilities(all) # probabilities of each classification
all.summary <- get_scpred(all) # summary table
```

## 8.
Testing model: % accuracy with a known-value test set

```{r}
testing <- merge(testing.cells, res.testing)
testing <- NormalizeData(testing)
all.test <- scPredict(testing, all)

all.verify <- crossTab(all.test, "phenotype", "scpred_prediction")
# 472/530 = 89.06% success
# columns = each type of classified neuron
# rows = where each neuron was classified
```

## 9.
ML model with global genes summary (boostrap 1)
· 89.06% success based on 472/530 testing data points
· res: ROC 0.992, sensitivity 0.957, specificity 0.96
· vuln: ROC 0.992, sensitivity 0.96, specificity 0.957

Iterating across 5 bootstrap replicates (resampling training and testing data):
· Average success rate:     88.303%
· Average ROC:              0.9916
· Average res sensitivity:  0.9566
· Average res selectivity:  0.9586
· Average vuln sensitivity: 0.9586
· Average vuln sensitivity: 0.9566
